\documentclass[xcolor=table,dvipsnames]{beamer}

\usepackage{lscape, amsmath, amsfonts, amssymb, setspace, theorem, wrapfig, graphicx, float, multirow, subfig, color, rotating, multicol, datetime, natbib, venndiagram, pstricks, xkeyval, tikz, etoolbox, verbatim}

\usepackage{listings}
\usepackage{xcolor}
 
\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegreengray}{rgb}{0,0.4,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codeblue}{rgb}{0.00,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}
\definecolor{jeopardy}{rgb}{.24,.47,.914}

\lstdefinestyle{mystyle}{
    backgroundcolor=\color{backcolour},   
    commentstyle=\color{codegreengray},
    numberstyle=\tiny\color{codegray},
    stringstyle=\color{codegreen},
    basicstyle=\ttfamily\footnotesize,
    breakatwhitespace=false,         
    breaklines=true,                 
    captionpos=b,                    
    keepspaces=true,                 
    numbers=left,                    
    numbersep=5pt,                  
    showspaces=false,                
    showstringspaces=false,
    showtabs=false,                  
    tabsize=2
}
 
\lstset{style=mystyle}

\title{GV300 - Quantitative Political Analysis}
\subtitle{University of Essex - Department of Government}
\date{Week 19 -- 3 February, 2020}				% or you can specify a date, just write it down instead of "\today"
\author{Lorenzo Crippa} 

\usetheme[progressbar=frametitle]{metropolis}
\usecolortheme{seahorse}						% try others: wolverine; crane...


\begin{document}
\frame{
\titlepage
}

%\begin{frame}
%\frametitle{What do you want to do?}
%\begin{itemize}
%\item We watch all episodes of Game of Thrones at 8x times the speed. We count the number of killings. Whoever creates a probability distribution that best corresponds to the distribution of the random variable ''number of death in GOT'' wins 5 extra marks on the last problem set. \pause
%\item We are randomly mixing karaoke, jeopardy, and Britain got talent for 5 extra marks on the last problem set. 
%\item We re-watch a Tottenham match after Mourinho took over and count the number of attempts on goal. Whoever creates a probability distribution that best corresponds to the distribution of the random variable ''number of attempts on goal'' wins 5 extra marks on the last problem set.
%\end{itemize}
%\end{frame}

%\begin{frame}
%\frametitle{What do you want to do?}
%\begin{itemize}
%\item We watch all episodes of Game of Thrones at 8x times the speed. We count the number of killings. Whoever creates a probability distribution that best corresponds to the distribution of the random variable ''number of death in GOT'' wins 5 extra marks on the last problem set. 
%\item We are randomly mixing karaoke, jeopardy, and Britain got talent for 5 extra marks on the last problem set. 
%\item We re-watch a Tottenham match after Mourinho took over and count the number of attempts on goal. Whoever creates a probability distribution \ldots yeah, the answer is it is rare-events model \ldots negative binomial or known by laymen as ''destruction of football''. 
%\end{itemize}
%\end{frame}

\begin{frame}
\frametitle{Let's play!}
\centering 
We are randomly mixing karaoke, jeopardy, and Britain got talent for 5 extra marks on the last problem set. 
\end{frame}


\begin{frame}
\frametitle{Rules}
\begin{enumerate}
\item We first create teams by random assignment. \href{https://www.randomlists.com/team-generator}{(click)}
\item You choose a category and a question (starting from the lowest available points question)
\item If you think you got the answer, run to project screen and touch the respective points field and yell ``heteroskedasticity'', then present your answer.
\item Any other team who thinks they got a better answer can now intervene by yelling ``homoskedasticity''
\item The group giving the best answer wins the points associated with the question
\item The team with the most points at the end wins 5 marks for each of its members applied to the next problem set
\end{enumerate}
\end{frame}


\begin{frame}
\frametitle{Next games}
\begin{table}[H]\centering
{\color{white} \begin{tabular}{|r|c|c|c|} \hline
\rowcolor{jeopardy} & Your data sucks & Your model sucks & Your graph sucks \\\hline
\rowcolor{jeopardy} 100 & & & \\\hline
\rowcolor{jeopardy} 200 & & & \\\hline
\rowcolor{jeopardy} 300 & &	& \\\hline
\end{tabular}}
\end{table}

\end{frame}	


\begin{frame}
\frametitle{Your Data sucks 100}
\begin{itemize}	
\item[] Fit a linear model on a dependent variable. Create a new variable and add it to the linear model. Create the variable in a way that it helps us as little as possible to learn about the relationship between dependent variable and that new variable (you got 5 mins).
\end{itemize}
\end{frame}	

\begin{frame}
\frametitle{Next games}
\begin{table}[H]\centering
{\color{white} \begin{tabular}{|r|c|c|c|} \hline
\rowcolor{jeopardy} & Your data sucks & Your model sucks & Your graph sucks \\\hline
\rowcolor{jeopardy} 100 & & & \\\hline
\rowcolor{jeopardy} 200 & & & \\\hline
\rowcolor{jeopardy} 300 & &	& \\\hline
\end{tabular}}
\end{table}

\end{frame}	


\begin{frame}
\frametitle{Your Data sucks 200}
\begin{itemize}	
\item[] Find a significant difference in an outcome variable separated by a grouping variable. Start omitting data. Find the smallest sample that still shows a significant difference (you got 5 mins).  
\end{itemize}
\end{frame}	

\begin{frame}
\frametitle{Next games}
\begin{table}[H]\centering
{\color{white} \begin{tabular}{|r|c|c|c|} \hline
\rowcolor{jeopardy} & Your data sucks & Your model sucks & Your graph sucks \\\hline
\rowcolor{jeopardy} 100 & & & \\\hline
\rowcolor{jeopardy} 200 & & & \\\hline
\rowcolor{jeopardy} 300 & &	& \\\hline
\end{tabular}}
\end{table}

\end{frame}	


\begin{frame}
\frametitle{Your Data sucks 300}
\begin{itemize}	
\item[] Fit an instrumental variable regression. Find the variable that (1) gives you the weakest instrument (2) and violates the exclusion restriction the most. (you got 5 mins).
\end{itemize}
\end{frame}	

\begin{frame}
\frametitle{Next games}
\begin{table}[H]\centering
{\color{white} \begin{tabular}{|r|c|c|c|} \hline
\rowcolor{jeopardy} & Your data sucks & Your model sucks & Your graph sucks \\\hline
\rowcolor{jeopardy} 100 & & & \\\hline
\rowcolor{jeopardy} 200 & & & \\\hline
\rowcolor{jeopardy} 300 & &	& \\\hline
\end{tabular}}
\end{table}

\end{frame}	


\begin{frame}
\frametitle{Your Model sucks 100}
\begin{itemize}	
\item[] Take a continuous variable as dependent variable and fit a linear model. Find the model with the highest $R^2$ (you got 5 mins). 
\end{itemize}
\end{frame}	

\begin{frame}
\frametitle{Next games}
\begin{table}[H]\centering
{\color{white} \begin{tabular}{|r|c|c|c|} \hline
\rowcolor{jeopardy} & Your data sucks & Your model sucks & Your graph sucks \\\hline
\rowcolor{jeopardy} 100 & & & \\\hline
\rowcolor{jeopardy} 200 & & & \\\hline
\rowcolor{jeopardy} 300 & &	& \\\hline
\end{tabular}}
\end{table}

\end{frame}	


\begin{frame}
\frametitle{Your Model sucks 200}
\begin{itemize}	
\item[] Take a binary dependent variable and fit a non-linear model. Find the model with the highest negative log-likelihood within (you got 5 mins). 
\end{itemize}
\end{frame}	

\begin{frame}
\frametitle{Next games}
\begin{table}[H]\centering
{\color{white} \begin{tabular}{|r|c|c|c|} \hline
\rowcolor{jeopardy} & Your data sucks & Your model sucks & Your graph sucks \\\hline
\rowcolor{jeopardy} 100 & & & \\\hline
\rowcolor{jeopardy} 200 & & & \\\hline
\rowcolor{jeopardy} 300 & &	& \\\hline
\end{tabular}}
\end{table}

\end{frame}	


\begin{frame}
\frametitle{Your Model sucks 300}
\begin{itemize}	
\item[] Fit a non-linear model. Find a model with the smallest number of variables that does not converge (you got 5 mins). 
\end{itemize}
\end{frame}	

\begin{frame}
\frametitle{Next games}
\begin{table}[H]\centering
{\color{white} \begin{tabular}{|r|c|c|c|} \hline
\rowcolor{jeopardy} & Your data sucks & Your model sucks & Your graph sucks \\\hline
\rowcolor{jeopardy} 100 & & & \\\hline
\rowcolor{jeopardy} 200 & & & \\\hline
\rowcolor{jeopardy} 300 & &	& \\\hline
\end{tabular}}
\end{table}

\end{frame}	


\begin{frame}
\frametitle{Your Graph sucks 100}
\begin{itemize}	
\item[] Show a binary relationship in a scatter plot. Find the best manipulation of the graph's layout to fake the visual existence of a substantial effect in the relationship (you got 5 mins). 
\end{itemize}
\end{frame}	

\begin{frame}
\frametitle{Next games}
\begin{table}[H]\centering
{\color{white} \begin{tabular}{|r|c|c|c|} \hline
\rowcolor{jeopardy} & Your data sucks & Your model sucks & Your graph sucks \\\hline
\rowcolor{jeopardy} 100 & & & \\\hline
\rowcolor{jeopardy} 200 & & & \\\hline
\rowcolor{jeopardy} 300 & &	& \\\hline
\end{tabular}}
\end{table}

\end{frame}	


\begin{frame}
\frametitle{Your Graph sucks 200}
\begin{itemize}	
\item[] Fit a linear model, get predicted values and plot against the dependent variable. Find the model that gives you the most perfect fit visually (you got 5 mins). 
\end{itemize}
\end{frame}	

\begin{frame}
\frametitle{Next games}
\begin{table}[H]\centering
{\color{white} \begin{tabular}{|r|c|c|c|} \hline
\rowcolor{jeopardy} & Your data sucks & Your model sucks & Your graph sucks \\\hline
\rowcolor{jeopardy} 100 & & & \\\hline
\rowcolor{jeopardy} 200 & & & \\\hline
\rowcolor{jeopardy} 300 & &	& \\\hline
\end{tabular}}
\end{table}

\end{frame}	


\begin{frame}
\frametitle{Your Graph sucks 300}
\begin{itemize}	
\item[] Fit a linear model, create a new variable that creates heteroskedasticity in the errors and add this variable to the model. Find the new variable that visually maximises the amount of heteroskedasticity (you got 5 mins).
\end{itemize}
\end{frame}	

\begin{frame}
\frametitle{Next games}
\begin{table}[H]\centering
{\color{white} \begin{tabular}{|r|c|c|c|} \hline
\rowcolor{jeopardy} & Your data sucks & Your model sucks & Your graph sucks \\\hline
\rowcolor{jeopardy} 100 & & & \\\hline
\rowcolor{jeopardy} 200 & & & \\\hline
\rowcolor{jeopardy} 300 & &	& \\\hline
\end{tabular}}
\end{table}

\end{frame}	


\end{document}